---
title: "01_Disturbance_Plot_Level"
format:
  html:
    theme: spacelab
    toc: true
    number-sections: true
---

# Overview

In this document, I will calculate canopy disturbance percentages from tree level up to plot level across the Balkan countries: Bosnia and Herzegovina, Albania, Bulgaria, and Croatia and the Semenic stand from Romania.

# Gather data & identifiers

## Packages

```{r}
#| echo: false
#| output: false

library(tidyverse);library(pool);library(zoo);library(pracma);library(lme4);library(cutpointr); library(pool); library(readxl); library(purrr); library(readr); library(forecast); library(dplR); library(ggrepel)

```

## Database Pull

**Pull from source that contains functions utilized in this code**

```{r}
source("B00_disturbance_functions.R")
census_1trees <- read_csv("Data/full_trees_dist.csv") %>%
  select(-1)
```

**Database access**

```{r}
# Connect to the database
KELuser <- dbPool(RPostgreSQL::PostgreSQL(),
                  dbname = 'remoteforestsorg',
                  host = '91.239.201.14',
                  port = 4010,
                  user = 'remoteforests002',
                  password = 'COVBtxQ5')
```

## Identifiers

::: panel-tabset
### Plot ID

```{r}
plot.id <- tbl(KELuser, "plot") %>%
  filter(ownership %in% 1,
          country %in% c("Albania", "Bosnia", "Croatia", "Romania", "Bulgaria"),
          foresttype %in% c("beech"),
          plottype %in% c(3, 4),
         census %in% 1,
         !is.na(lng),
         !is.na(lat)) %>%
 filter(!location %in% c("Fagaras", "Maramures")) %>%
pull(id)
```

### Tree ID

```{r}
tree.id <- census_1trees %>% 
  filter(dbh_mm >= 60,
         !status %in% c(0, 10, 15, 99),
         growth %in% c(-1, 1, 99),
         treetype %in% "0" & onplot %in% c(1, 2) | treetype %in% c("m", "x"),
         !species %in% c("Lians", "99")) %>%
  pull(tree_id)

 # tree.id <- tbl(KELuser, "tree") %>%
 #  filter(dbh_mm >= 60,
 #         !status %in% c(0, 10, 15, 99),
 #         #growth %in% c(-1, 1, 99),
 #         treetype %in% "0" & onplot %in% c(1, 2) | treetype %in% c("m", "x"),
 #         !species %in% c("Lians", "99")) %>%
 #  collect()
 # 
 
```


### Core ID

```{r}
core.id <- tbl(KELuser, "core") %>%
  filter(coretype %in% c(1,2),
         !crossdated %in% c(12, 20:22, 99),
         !corestatus %in% c(2, 3),
         !is.na(missing_mm),
         !is.na(missing_years),
         missing_mm < 30,
         missing_years < 30) %>%
  pull(id)
```

:::

# Tree-level disturbance

**Create data list**


```{r}
ID <- tbl(KELuser, "core") %>% filter(id %in% core.id) %>%
  inner_join(., tbl(KELuser, "tree") %>% filter(id %in% tree.id), by = c("tree_id" = "id")) %>%
  inner_join(., tbl(KELuser, "plot") %>% filter(id %in% plot.id), by = c("plot_id" = "id")) %>%
  pull(id)

data.list <- distGetData(ID = ID)
```

## Growth reconstruction

```{r}
data.growth <- growthCalculate(data = data.list, windowLength = 10)
```


## Identify events - gap ascension

```{r}
data.event <- eventCalculate(data = data.growth, gapAge = c(5:15), nprol = 7) #3772
```

# Plot level

## Crown area

Crown area is constructed through a linear model whereby canopy area is predicted by dbh_mm taking stand, subplot, and plot into account. There is an individual model for coniferous and broadleaf species.

```{r}
data.ca <- tbl(KELuser, "dist_data_ca") %>% select(stand, subplot, plotid, sp_type, dbh_mm, ca_m2) %>% collect()

data.conif <- data.ca %>% filter(sp_type %in% "coniferous")
conif <- lmer(sqrt(ca_m2) ~ dbh_mm + (1|stand) + (1|subplot) + (1|plotid), data = data.conif)

data.broad <- data.ca %>% filter(sp_type %in% "broadleaved")
broad <- lmer(sqrt(ca_m2) ~ dbh_mm + (1|stand) + (1|subplot) + (1|plotid), data = data.broad)
```

## Sampled & unsampled

::: panel-tabset
### Sampled trees

Sampled trees contain those on the plot that were cored and identified as usable cores (adequate correlation to master, minimal missing rings, etc.). I also filtered out plots where there was not a minimum of 8 usable trees per plot. This number is considered a minimum sample size for cores in a plot (really should be higher like 10 or 15).

```{r}
data.sampled <- data.event %>% 
  inner_join(., tbl(KELuser, "core") %>% select(core_id = id, tree_id), by = "tree_id", copy = TRUE) %>%
  inner_join(., tbl(KELuser, "ring") %>% 
               group_by(core_id) %>% summarise(year_min = min(year), year_max = max(year) - 10),
             by = "core_id", copy = TRUE) %>%
  filter(!(year_min > year_max & year > year_max)) %>%
  inner_join(., tbl(KELuser, "tree") %>% filter(plot_id %in% plot.id), by = c("tree_id" = "id"), copy = TRUE) %>%
  distinct(., plot_id, tree_id, year_min, year_max) %>%
  group_by(plot_id) %>%
  # filter plots by the minimum number of usable cores
  filter(n() >= 8) %>%
  ungroup() %>%
  inner_join(., tbl(KELuser, "plot") %>% select(-stand, -subplot), by = c("plot_id" = "id"), copy = TRUE) %>%
  inner_join(., tbl(KELuser, "spatial_hierarchy"), by = "plotid", copy = TRUE) %>%
  inner_join(., tbl(KELuser, "tree"), by = c("tree_id" = "id", "plot_id"), copy = TRUE) %>%
  inner_join(., tbl(KELuser, "species_fk"), by = c("species" = "id"), copy = TRUE) %>%
  inner_join(., data.event, by = "tree_id") %>%
  select(country, stand, subplot, plotid, growth, treeid, tree_id, sp_type, species, dbh_mm, year_min, year_max, event, year) %>%
  collect()


# data.sampled %>%
#   ggplot()+
#   geom_point(aes())


data.sampled %>%
  distinct(., plotid, .keep_all = TRUE) %>%
  group_by(stand) %>%
 summarise(n = n())
```

### Unsampled trees

The unsampled trees are not cored or have unusable cores and are located on the plot - should also be canopy level trees.

```{r}
data.unsampled <- tbl(KELuser, "core") %>%
  filter(coretype %in% 1) %>%
  collect() %>%
  right_join(., census_1trees %>% 
               select(tree_id, treeid, growth, status, layer, species, dbh_mm, plot_id) %>%
               filter(growth %in% c(-1, 1, 99),
                    !tree_id %in% local(data.sampled$tree_id)), by = c("tree_id" = "tree_id")) %>% 
  inner_join(., tbl(KELuser, "tree") %>% filter(id %in% local(data.sampled$tree_id)) %>% distinct(., plot_id), by = "plot_id", copy = TRUE) %>%
  inner_join(., tbl(KELuser, "plot"), by = c("plot_id" = "id"), copy = TRUE) %>%
  filter((foresttype %in% "spruce" & !stand %in% "Polana" & !is.na(subcore)) | ((foresttype %in% "beech" | stand %in% "Polana") & (status %in% c(1:4) & !is.na(subcore)))) %>%
  select(-stand, -subplot) %>%
  inner_join(., tbl(KELuser, "spatial_hierarchy"), by = "plotid", copy = TRUE) %>%
  inner_join(., tbl(KELuser, "species_fk"), by = c("species" = "id"), copy = TRUE) %>%
  mutate(year_min = 0,
         year_max = 0,
         event = "unsampled",
         year = 0) %>%
  select(country, stand, subplot, plotid, growth, treeid, tree_id, sp_type, species, dbh_mm, year_min, year_max, event, year) %>%
  collect()

data.unsampled %>%
  distinct(., plotid, tree_id, .keep_all = TRUE) %>%
  group_by(plotid) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

### Combined & canopy area calculation

```{r}
data.all <- bind_rows(data.sampled, data.unsampled) %>%
  mutate(ca_m2 = case_when(
          sp_type %in% "broadleaved" ~ predict(object = broad, newdata = ., allow.new.levels = T),
          sp_type %in% "coniferous" ~  predict(object = conif, newdata = ., allow.new.levels = T)),
         ca_m2 = ca_m2^2)

write_csv(data.all,file="Data/data.all.trees.csv") 

data.all<-read_csv("Data/data.all.trees.csv")

 # minimum common number of tree records per plot (sampled + unsampled)
data.all %>%
  distinct(., plotid, tree_id, .keep_all = TRUE) %>%
  group_by(plotid) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```


:::

# Bootstrapping plot-level

::: panel-tabset
## Calculate canopy disturbance percentage

```{r}
#| output: false

data.boot <- data.all %>% 
  distinct(., plotid, tree_id) %>%
  slice(rep(1:nrow(.), each = 1000)) %>%
  mutate(rep = rep(1:1000, times = nrow(.) / 1000)) 

set.seed(1)

data.dist.boot <- data.boot %>%
  group_by(plotid, rep) %>%
  # size of bootstrapping subsamples = maximum common number of tree records per plot (sampled + unsampled) - raise overlap??? 
  slice_sample(., n = 12, replace = TRUE) %>%
  ungroup() %>%
  left_join(., data.all %>% select(plotid, tree_id, year_min, year_max, event, year, ca_m2), by = c("plotid", "tree_id")) %>%
  do({
    
    x <- .
    
    x1 <- x %>% 
      filter(!event %in% c("unsampled")) %>% 
      group_by(plotid, rep, year) %>% 
      summarise(ca_event = sum(ca_m2)) %>%
      ungroup()
    
    x2 <- x %>% 
      group_by(plotid, rep, tree_id) %>% 
      filter(year %in% first(year)) %>% 
      group_by(plotid, rep) %>% 
      summarise(ca_total = sum(ca_m2)) %>%
      ungroup()
    
    x3 <- x %>% 
      group_by(plotid, rep, tree_id) %>% 
      filter(!event %in% "unsampled", 
             year %in% first(year)) %>% 
      group_by(plotid, rep) %>% 
      filter(n() >= 1) %>% 
      summarise(year_min = min(year_min),
                year_max = max(year_max),
                ncores = n()) %>%
      ungroup()
    
    inner_join(x1, x2 %>% inner_join(., x3, by = c("plotid", "rep")), by = c("plotid", "rep"))
  }) %>%
  mutate(ca_pct = ca_event * 100 / ca_total)
```


## Fill the period

```{r}
#| echo: false

beg <- min(data.dist.boot$year)
end <- max(data.dist.boot$year)

data.dist <- tibble()

for (p in unique(data.dist.boot$plotid)) {
  
  x <- data.dist.boot %>% filter(plotid %in% p) 
  
  out <- inner_join(
    x %>% 
      select(plotid, rep, year, ca_pct) %>% 
      group_by(plotid, rep) %>% 
      complete(year = beg:end, fill = list(ca_pct = 0)) %>% 
      group_by(plotid, year) %>% 
      summarise(ca_pct = mean(ca_pct)) %>%
      ungroup(),
    x %>% 
      distinct(., plotid, rep, year_min, year_max, ncores) %>% 
      group_by(plotid) %>%
      summarise(year_min = round(mean(year_min), 0),
                year_max = round(mean(year_max), 0),
                ncores = round(mean(ncores), 0)) %>%
      ungroup(),
    by = "plotid") 
  
  data.dist <- bind_rows(data.dist, out)
  
  remove(x, out)
}
```
:::

# Kernel density estimation & peak detection

::: panel-tabset
## Full chronologies

```{r}
data.kde.full <- data.dist %>%
  filter(year >= year_min, year <= year_max) %>%
  group_by(plotid) %>%
  complete(year = (min(year)-30):(max(year)+30), fill = list(ca_pct = 0)) %>%
  mutate(kde = kdeFun(ca_pct, k = 30, bw = 5, st = 7)) %>%
  filter(year %in% c((min(year)+15):(max(year)-15))) %>%
  ungroup() %>%
  mutate(type = "full")

data.peaks.full <- data.kde.full %>% #446
  group_by(plotid) %>%
  filter(row_number() %in% peakDetection(x = kde, threshold = 10, nups = 5, mindist = 10)) %>%
  ungroup() %>%
  mutate(type = "full",
         peak = "yes")
```

:::

# Plot disturbance at the plot level

::: panel-tabset
## Prep data

Below I plot the data for the data peaks.

```{r}
plot_level <- data.kde.full %>%
  drop_na() %>%
  left_join(data.peaks.full %>% select(peak, plotid, year), by = c("plotid", "year")) %>%
  mutate(peak = ifelse(peak %in% "yes", "yes", "no")) %>%
  left_join(data.sampled %>% distinct(., stand, plotid), by = "plotid")  %>%
  group_by(plotid) %>%
  filter(!year < year_min, 
         !year > year_max)
```


## Plot-level Disturbance

```{r}
rec_dis <- plot_level %>%
  filter(peak %in% "yes") %>%
  group_by(stand, plotid) %>%
  arrange(desc(year)) %>%
  summarise(my_rec_year = first(year),
            my_rec_sev = first(kde)) %>%
  ungroup()

max_dis <- plot_level %>%
  filter(peak %in% "yes") %>%
  group_by(stand, plotid) %>%
  arrange(desc(kde)) %>%
  summarise(my_max_year = first(year),
            my_max_sev = first(kde)) %>%
  ungroup()

n_peak <- plot_level %>%
    filter(peak %in% "yes") %>%
  group_by(plotid) %>%
  summarise(n_peak = n())

decadal_mean<-plot_level %>%
  mutate(decade = floor(year/10)*10) %>%
  group_by(stand, plotid, decade) %>%
  summarise(ca_pct_sum = sum(ca_pct),
            kde_mn = mean(kde),
            kde_sd = sd(kde)) %>%
  ungroup() %>%
  group_by(stand, plotid) %>%
  summarise(decade_mn = mean(kde_mn),
            # kde_sd = sd(kde_sd),
            n_decade = n()) %>%
  arrange(n_decade)

dist_joined <-  max_dis %>%
  inner_join(., rec_dis, by = c("stand", "plotid")) %>%
  inner_join(., decadal_mean, by = c("stand", "plotid")) %>%
  inner_join(.,  n_peak, by = "plotid")

remove(rec_dis, n_peak, max_dis, decadal_mean)

write_csv(dist_joined,file="Data/disturbance.csv")
```

:::

# Close database connection

```{r}
poolClose(KELuser)
```

